---
title: "Untitled"
author: "TK Wallace"
date: "2023-05-01"
output: html_document
---

```{r libraries and packages, include=FALSE, e}
#Install Packages
install.packages("Rtools")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages('stringr')
install.packages("installr")
install.packages("dplyr")
install.packages("gtable")
install.packages("cooccur")
install.packages("installr")
install.packages("snowfall")
install.packages("parallel")
install.packages("colorspace")
install.packages("latticeExtra")
install.packages("ggfortify")
install.packages("rnaturalearth")
install.packages("tidygeocoder")
install.packages("gcKrig")
#Libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(installr)
library(gtable)
library(purrr)
library(lubridate)
library(cooccur)
library(installr)
library(akima)
library(geoR) 
library(maps)
library(maptools)
library(lattice)
library(sp)
library(gstat)
library(stats)
library(corrplot)
library(stats)
library(snowfall)
library(gcKrig)
library(parallel)
library(colorspace)
library(latticeExtra)
library(mapdata)
library(ggplot2)
library(ggfortify)
library(sf)
library(ggspatial)
library(rnaturalearth)
library(tidygeocoder)
library(maps)
library(usmap)
```

## An analysis of Pathogens in Ixodes scapularis nymphs in Rhode Island

the code and generated tables, graphs, and outputs are an extensive analysis of pathogens present within Ixodes scapularis nymphs collected in Rhode Island during the summer (May, June, July) of 2022. We attempt to understand general pathogen prevalence for each of our tested pathogens, environmental associations with each pathogen, associations between pathogens at the site and individual level, and the spatial distribution of infected nymphs by pathogen

```{r imported data sheets, include=FALSE}
#Read Tables and import
pathogen_data_nymphs_2022 <- read_csv("pathogen_data_nymphs_2022_Enhanced.csv", 
                                      locale=locale(encoding="latin1")) #this table just has the specimen and pathogen informations

Sampling_Site_Longitude_and_Lattitude <- read_csv("Sampling Site Longitude and Lattitude.csv") #longitude and lattitude

Specimen_Date_Temp_Humidity <- read_csv("Specimen_Date_Temp_Humidity.csv") #temp and humidity data for each collection event

Sampling_Site_Enviromental_Parameter <- read_csv("Sampling Site Enviromental Parameter.csv") # TAMES and % Forest Cover

```

make sure that before you begin this analyses that you have these 4 data sheets within your local project or directory. Each contains different and unique information denoted by comments (Long, lat, enviromental conditions, collector informations, etc)

```{r imported data sheets, include=FALSE}
#Modify Specimen environmental and collection data for ease of use
Specimen_Date_Temp_Humidity_corrected <- Specimen_Date_Temp_Humidity %>%
  separate(Date, sep='_', into =c("month","day","year")) %>%
  rename(Site_Code = `SITE CODE`) %>%
  rename(Vial_code = `VIAL CODE`)

#convert vial code column to character to allow for join later (Join wont happen unless columns are same class)
Specimen_Date_Temp_Humidity_corrected$Vial_code <- as.character(
  Specimen_Date_Temp_Humidity_corrected$Vial_code)

#Change All Instances Of Positive and Negative To 1 and 0 for analysis
pathogen_data_nymphs_2022[pathogen_data_nymphs_2022=="Positive"] <- "1" 

pathogen_data_nymphs_2022[pathogen_data_nymphs_2022=="Negative"] <- "0"

#Remove all instances of non-ixodes and failed DNA extraction
#additionally, only include pathogens that are present

Pathogen_data_Nofails_OnlyIxodes <- pathogen_data_nymphs_2022 %>%
  filter(`Tick DNA Internal quality`=='Pass') %>%
  filter(`Dermacentor variabilis`==0) %>%
  filter(`Amblyomma americanum`==0) %>%
  select(c(`URI_Nymph_Number_on_Tube`,
           `URI_Vial_Number_on_tube`,
           'Ixodes scapularis',
           `Borrelia Genus`,
           `Borrelia burgdorferi sensu lato`,
           `Borrelia miyamotoi `,
           `Borrelia mayonii `,
           `Babesia microti`,
           `Ehrlichia muris`,
           `Anaplasma phagocytophilum`,
           `Deer Tick Virus`))

#Convert Pathogen Columns to Numeric for calcualation

Pathogen_data_numeric <- Pathogen_data_Nofails_OnlyIxodes %>%
  mutate_at(c('Borrelia Genus',
              'Borrelia burgdorferi sensu lato',
              'Borrelia miyamotoi ',
              'Borrelia mayonii ',
              'Babesia microti',
              'Ehrlichia muris',
              'Anaplasma phagocytophilum',
              'Deer Tick Virus',
              'Ixodes scapularis'), as.numeric)

#This creates a column in pathogen_data_numeric summarizing co-infections based on presence and infection amount

Pathogen_data_numeric$infection_amount <- rowSums(Pathogen_data_numeric[,5:11])

Pathogen_data_numeric <- Pathogen_data_numeric %>%
  mutate(Co_infection_present = if_else(infection_amount > 1,1,0))

#convert Vial and Specimen Number to character
Pathogen_data_numeric_pain <- Pathogen_data_numeric %>%
  mutate_at(c('URI_Nymph_Number_on_Tube', 'URI_Vial_Number_on_tube'), as.character) %>%
  mutate(Infection_present =
          if_else(`Borrelia Genus` | `Borrelia burgdorferi sensu lato` 
                  | `Borrelia miyamotoi ` | `Borrelia mayonii ` | `Babesia microti` 
                  | `Ehrlichia muris` | `Anaplasma phagocytophilum` | `Deer Tick Virus` == 1,1,0)) %>%
relocate(Infection_present, .after = `Ixodes scapularis`) 

Pathogen_data_numeric_pain <- Pathogen_data_numeric_pain[Pathogen_data_numeric_pain$`Ixodes scapularis` !=0, ]

#All of the above code will give us information on on individual nymph pathogen information

#_______________________________________________________________________________

##All of the code below this will help us summarise information at the site and time level

#Summarize total infected by vial number

Pathogen_data_total_infected_by_site <- Pathogen_data_numeric_pain %>%
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_infected_Ixodes = sum(`Infection_present`))
  

###Summarize by total positive specimens
  
#Borrelia burgdorferi
Pathogen_data_Total_Borrelia_burgdorferi <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Borrelia_burgdorferi = sum(`Borrelia burgdorferi sensu lato`))
 
#Ixodes Scapularis
Pathogen_data_Total_Ixodes_scapularis <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Ixodes_Scapularis = sum(`Ixodes scapularis`))

#Borrelia genus
Pathogen_data_Total_Borrelia_genus <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Borrelia_genus= sum(`Borrelia Genus`))

#Borrelia miyamotoi
Pathogen_data_Total_Borrelia_miyamotoi <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Borrelia_miyamotoi= sum(`Borrelia miyamotoi `))

#Borrelia Mayonii
Pathogen_data_Total_Borrelia_mayonii <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube) %>%
  summarise(total_Borrelia_Mayonii = sum(`Borrelia mayonii `))

#Babesia Microti
Pathogen_data_Total_microti <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Babesia_microti = sum(`Babesia microti`))

#Ehrlichia Muris
Pathogen_data_Total_ehrlichia <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Ehrlichia_muris = sum(`Ehrlichia muris`))

#Anaplasma Phagoctyophilum
Pathogen_data_Total_phagocytophilum <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Anaplasma_phagocytophilum = sum(`Anaplasma phagocytophilum`))

#Deer Tick Virus
Pathogen_data_Total_Powassan_DTV <- Pathogen_data_numeric_pain %>% 
  group_by(URI_Vial_Number_on_tube)  %>%
  summarise(total_Powassan_DTV = sum(`Deer Tick Virus`))

##make sure you use Co_infection_presents column, that column specifies is co-infections are present but not if it is a triple or secondary infection
# Number of Co-infections
Pathogen_data_total_coinfections <- Pathogen_data_numeric_pain %>%
  group_by(URI_Vial_Number_on_tube) %>%
  summarise(total_coinfected_individuals = sum(`Co_infection_present`))
            
#Merge all into single data
Pathogen_list <- list(Pathogen_data_Total_Ixodes_scapularis,
                      Pathogen_data_Total_Borrelia_burgdorferi,
                      Pathogen_data_Total_Borrelia_genus,
                      Pathogen_data_Total_Borrelia_miyamotoi,
                      Pathogen_data_Total_Borrelia_mayonii,
                      Pathogen_data_Total_microti,
                      Pathogen_data_Total_ehrlichia,
                      Pathogen_data_Total_phagocytophilum,
                      Pathogen_data_Total_Powassan_DTV,
                      Pathogen_data_total_infected_by_site,
                      Pathogen_data_total_coinfections)

Pathogen_combined_full <- Pathogen_list %>%
  reduce(full_join, by = 'URI_Vial_Number_on_tube') %>%
  relocate(total_infected_Ixodes, .after = total_Ixodes_Scapularis)

#Reorganize columns 
Pathogen_with_sites_codes_long_lat <- Pathogen_combined_full %>%
  rename(Vial_code = URI_Vial_Number_on_tube) %>%
  select(Vial_code:total_coinfected_individuals) 
 
Pathogen_with_sites_codes_and_all_information <- Pathogen_with_sites_codes_long_lat %>%
  left_join(Specimen_Date_Temp_Humidity_corrected, by = ('Vial_code')) %>%
  left_join(Sampling_Site_Longitude_and_Lattitude, by = ('Site_Code')) %>%
  relocate(Site_Code, .after = Vial_code) %>%
  relocate(`Longitude (x) -`, .after = Site_Code) %>%
  relocate(`Latitude (y) +`, .after = `Longitude (x) -`)


#Pathogen all by site
#Additionally, it provides a percentage of ticks at each site that are infected with pathogen

Pathogen_by_site_with_percentages <- Pathogen_with_sites_codes_and_all_information %>%
  select(c(Site_Code:Humdity)) %>%
  group_by(Site_Code) %>%
  summarise(total_Ixodes_Scapularis_nymphs = sum(total_Ixodes_Scapularis),
            total_infected_ixodes_scapularis_nymphs = sum(total_infected_Ixodes),
            total_Borrelia_burgdorferi = sum(total_Borrelia_burgdorferi),
            total_Borrelia_Mayonii = sum(total_Borrelia_Mayonii),
            total_Borrelia_miyamotoi = sum(total_Borrelia_miyamotoi),
            total_Babesia_microti = sum(total_Babesia_microti),
            total_Ehrlichia_muris = sum(total_Ehrlichia_muris),
            total_Anaplasma_phagocytophilum = sum(total_Anaplasma_phagocytophilum),
            total_Powassan_DTV = sum(total_Powassan_DTV),
            total_coinfection_individuals = sum(total_coinfected_individuals)) %>%
  mutate(percent_infected_nymphs = total_infected_ixodes_scapularis_nymphs/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Bb = total_Borrelia_burgdorferi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_miyamotoi = total_Borrelia_miyamotoi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_mayonii = total_Borrelia_Mayonii/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Babesia_microti = total_Babesia_microti/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Ehrlichia_muris = total_Ehrlichia_muris/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Anaplasma_phagocytophilum = total_Anaplasma_phagocytophilum/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Powassan_DTV = total_Powassan_DTV/total_Ixodes_Scapularis_nymphs*100,
         percent_total_coinfection_individuals = total_coinfection_individuals/total_Ixodes_Scapularis_nymphs*100)

#Pathogens all by month
#Additionally, it provides a percentage of ticks at each site that are infected with pathogen

Pathogen_by_month_with_percentages <- Pathogen_with_sites_codes_and_all_information %>%
  select(c(Vial_code:Humdity)) %>%
  group_by(month) %>%
  summarise(total_Ixodes_Scapularis_nymphs = sum(total_Ixodes_Scapularis),
            total_infected_ixodes_scapularis_nymphs = sum(total_infected_Ixodes),
            total_Borrelia_burgdorferi = sum(total_Borrelia_burgdorferi),
            total_Borrelia_Mayonii = sum(total_Borrelia_Mayonii),
            total_Borrelia_miyamotoi = sum(total_Borrelia_miyamotoi),
            total_Babesia_microti = sum(total_Babesia_microti),
            total_Ehrlichia_muris = sum(total_Ehrlichia_muris),
            total_Anaplasma_phagocytophilum = sum(total_Anaplasma_phagocytophilum),
            total_Powassan_DTV = sum(total_Powassan_DTV),
            total_coinfection_individuals = sum(total_coinfected_individuals)) %>%
  mutate(percent_infected_nymphs = total_infected_ixodes_scapularis_nymphs/total_Ixodes_Scapularis_nymphs*100,
    percent_infection_Bb = total_Borrelia_burgdorferi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_miyamotoi = total_Borrelia_miyamotoi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_mayonii = total_Borrelia_Mayonii/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Babesia_microti = total_Babesia_microti/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Ehrlichia_muris = total_Ehrlichia_muris/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Anaplasma_phagocytophilum = total_Anaplasma_phagocytophilum/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Powassan_DTV = total_Powassan_DTV/total_Ixodes_Scapularis_nymphs*100,
    percent_total_coinfection_individuals = total_coinfection_individuals/total_Ixodes_Scapularis_nymphs*100)


#Pathogens by month and by site
#Additionally, it provides a percentage of ticks at each site that are infected with pathogen

Pathogen_by_site_and_month_with_percentages <- Pathogen_with_sites_codes_and_all_information %>%
  select(c(Site_Code:Humdity)) %>%
  group_by(Site_Code, month) %>%
  summarise(total_Ixodes_Scapularis_nymphs = sum(total_Ixodes_Scapularis),
            total_infected_ixodes_scapularis_nymphs = sum(total_infected_Ixodes),
            total_Borrelia_burgdorferi = sum(total_Borrelia_burgdorferi),
            total_Borrelia_Mayonii = sum(total_Borrelia_Mayonii),
            total_Borrelia_miyamotoi = sum(total_Borrelia_miyamotoi),
            total_Babesia_microti = sum(total_Babesia_microti),
            total_Ehrlichia_muris = sum(total_Ehrlichia_muris),
            total_Anaplasma_phagocytophilum = sum(total_Anaplasma_phagocytophilum),
            total_Powassan_DTV = sum(total_Powassan_DTV),
            total_coinfection_individuals = sum(total_coinfected_individuals)) %>%
  mutate(percent_infected_nymphs = total_infected_ixodes_scapularis_nymphs/total_Ixodes_Scapularis_nymphs*100,
    percent_infection_Bb = total_Borrelia_burgdorferi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_miyamotoi = total_Borrelia_miyamotoi/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Borrelia_mayonii = total_Borrelia_Mayonii/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Babesia_microti = total_Babesia_microti/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Ehrlichia_muris = total_Ehrlichia_muris/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Anaplasma_phagocytophilum = total_Anaplasma_phagocytophilum/total_Ixodes_Scapularis_nymphs*100,
         percent_infection_Powassan_DTV = total_Powassan_DTV/total_Ixodes_Scapularis_nymphs*100,
    percent_total_coinfection_individuals = total_coinfection_individuals/total_Ixodes_Scapularis_nymphs*100)

#Join with environmental covariates
#can't do time here, you can only do site level

Pathogen_by_site_and_month_with_percentages <- Pathogen_by_site_and_month_with_percentages %>%
  left_join(Sampling_Site_Enviromental_Parameter, by = "Site_Code") 

  
Pathogen_by_site_with_percentages <- Pathogen_by_site_with_percentages %>%
  left_join(Sampling_Site_Enviromental_Parameter, by = "Site_Code")

#Relocate and select data.frames to make them more readable
Pathogen_by_site_and_month_with_percentages <- Pathogen_by_site_and_month_with_percentages %>%
  relocate(`Longitude (x) -`, .after = month) %>%
  relocate(`Latitude (y) +`, .after = `Longitude (x) -`)

Pathogen_by_site_with_percentages <- Pathogen_by_site_with_percentages %>%
  relocate(`Longitude (x) -`, .after = Site_Code) %>%
  relocate(`Latitude (y) +`, .after = `Longitude (x) -`)

#make a version of the site level counts with the buffers so it is only % forest cover for each site
#we will use this data frame for when we need to analyze covariates at the site level

Pathogen_by_site_spatial_analysis <- Pathogen_by_site_with_percentages %>%
  select(-c("percent_infected_nymphs", "percent_infection_Bb", "percent_infection_Borrelia_miyamotoi",
            "percent_infection_Borrelia_mayonii", "percent_infection_Babesia_microti","percent_infection_Ehrlichia_muris",
            "percent_infection_Anaplasma_phagocytophilum", "percent_infection_Powassan_DTV", "percent_total_coinfection_individuals",
  ))

Pathogen_by_site_spatial_analysis <- Pathogen_by_site_spatial_analysis %>%
  select(1:25) %>%
  select(-c("Site_Code"))

#only select columns we need
Pathogen_by_site_spatial_analysisz_no_buffers <- Pathogen_by_site_spatial_analysis %>%
  select(1,2,3,4,5,6,7,8,9,10,11,12,13,15,18,21,24)

#replace NA's with zeroes
Pathogen_by_site_spatial_analysisz_no_buffers[is.na(Pathogen_by_site_spatial_analysisz_no_buffers)] <- 0

```

```{r exporting data frames, include=FALSE}

#This section exports all previously collected data frames as .csv's

#CSV BY SITE regardless of month
write.csv(Pathogen_by_site_with_percentages, "../../Pathogen_by_site")

#Csv of month regardless of site
write.csv(Pathogen_by_month_with_percentages, "../../Pathogen_by_month_with_percentages")

#Csv of site and month
write.csv(Pathogen_by_site_and_month_with_percentages, "../../Pathogen_by_site_and_month_with_percentages")

```

After running the above code you should be able to simply export the information into csv's. Reminder that the data frames you generate are summary tables that generalize the information according to site, month, or both. The data the only includes individuals with enviromental covariates is  exported later within this document

## Exploratory Analysis of Pathogen Infection In Nymphs

Let's do some exploratory analysis but what does that entail. For this we will look at how many nymphs are infected with each pathogen and overall infection rate. Additionally, how many have co-infections compared to singular infections and what pathogen combinations makes up those co-infections

```{r Summing Up pathogen distribution, include=FALSE}
sum(Pathogen_by_month_with_percentages$total_Ixodes_Scapularis_nymphs) # 1566 amount
sum(Pathogen_by_month_with_percentages$total_infected_ixodes_scapularis_nymphs) #432 amount
sum(Pathogen_by_month_with_percentages$total_Borrelia_burgdorferi) #266 amount
sum(Pathogen_by_month_with_percentages$total_Borrelia_Mayonii) #0 amount
sum(Pathogen_by_month_with_percentages$total_Borrelia_miyamotoi) #14 amount
sum(Pathogen_by_month_with_percentages$total_Babesia_microti) #107 amount
sum(Pathogen_by_month_with_percentages$total_Ehrlichia_muris) #0 amount
sum(Pathogen_by_month_with_percentages$total_Anaplasma_phagocytophilum) #105 amount
sum(Pathogen_by_month_with_percentages$total_Powassan_DTV) #4 amount
sum(Pathogen_by_month_with_percentages$total_coinfection_individuals) #65 amount

Pathogen_by_month_bar_graph_df <- Pathogen_by_month_with_percentages %>%
  select(1,4,5,6,7,8,9,10)

#pathogen distribution by monthly prevalence

Pathogen_by_month_bar_graph_df <- Pathogen_by_month_with_percentages %>%
  rename(`Borrelia burgdorferi` = total_Borrelia_burgdorferi, 
         `Borrelia miyamotoi` = total_Borrelia_miyamotoi,
         `Babesia microti` = total_Babesia_microti,
         ` Anaplasma phagocytophilum`= total_Anaplasma_phagocytophilum,
         `Deer Tick Virus` = total_Powassan_DTV,
         `Borrelia Mayonii` = total_Borrelia_Mayonii,
         `Ehrlichia muris` = total_Ehrlichia_muris) %>%
  select(1,4,5,6,7,8,9,10)

Pathogen_by_month_bar_graph_df <- Pathogen_by_month_bar_graph_df %>%
  pivot_longer(!month, names_to = "Pathogen", values_to ="count")

Pathogen_by_month_bar_graph <- ggplot(Pathogen_by_month_bar_graph_df, aes(x = month, y = count, fill = Pathogen)) +
  geom_bar(stat="identity",position = "fill") 

Pathogen_by_month_bar_graph <- Pathogen_by_month_bar_graph + theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  panel.grid.major = element_blank(), #remove major gridlines
  panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(fill='transparent') #transparent legend panel
)

Pathogen_by_month_bar_graph <- Pathogen_by_month_bar_graph + ggtitle("% of Collected Nymphs Infected With Pathogen") +
  xlab("Month") + ylab("% of I. scapularis Nymphs") +
  scale_x_discrete(breaks=c("5","6","7"),
                   labels=c("May", "June", "July"))


```

We collected 1566 I. scapularis nymphs. Among those 432 were infected with at least 1 pathogen, 266 with Borrelia burgdorferi, 107 with Babesia microti, 105 with A. phagocytophilum, 14 with B. miyamotoi, and 4 with Powassan Virus/DTV. 65 of tested nymphs were co-infected (had 2 or more pathogens present simultaneously)

```{r plotting pathogen prevalence, ech=FALSE}
plot(Pathogen_by_month_bar_graph)
```
Above is a graph detailing pathogen prevalences across every month. Note that we did not sample every site, every month. For example, we sampled ourn sites for the first round across May and June, while the second round was done entirely in July. Given this i feel comfortable in saying that so far, we cannot make any strong inferences on the pheneology of pathogen prevalence for nymphs in Rhode Island. Currently, it seems that prevalence is constant across the nymphal season.

Now Let's see rates of co-infection across all infection categories

```{r infection status, include = FALSE}
#Now lets see what pathogens make up those co-infections (we need to use the raw by individual data)
infection_amount_table <- table(Pathogen_data_numeric_pain$infection_amount)
# 1134 No Infection
# 367 singular infection
# 63 double infection
# 2 triple infection
```
```{r infection status graph, include=FALSE}
nfection_amount_df <- as.data.frame(infection_amount_table)

infection_amount_bar_graph <- ggplot(infection_amount_df, aes(x = Var1, y = Freq)) +
  geom_bar(stat="identity", fill = "black") 

infection_amount_bar_graph <- infection_amount_bar_graph + theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  panel.grid.major = element_blank(), #remove major gridlines
  panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(fill='transparent') #transparent legend panel
)

infection_amount_bar_graph <- infection_amount_bar_graph + ggtitle("Frequency of Infection in I. scapularis nymphs") +
  xlab("Infection Type") + ylab("# of I. scapularis Nymphs") +
  scale_x_discrete(breaks=c("0","1","2","3"),
                   labels=c("No Infection",
                            "Single Infection",
                            "Double Infection",
                            "Triple Infection"))
```
```{r infection graph, echo=FALSE}
plot(infection_amount_bar_graph)
```
Across all nymphs 1134 nymphs were not infected with any pathogen, 367 had a singular infection, 63 had a double infection, and only 2 nymphs had a triple infection. Since Double infections are so common compared to triple infections, lets see which combinations in pathogens are more common than others.

```{r co-infection graph}
# triple infections, this was done by hand but if triple infections are high do what i did for double infections below
#but replace the infection_amount === 2 with infection_amount === 3 in the subset function

#BB/Bam/Ap = 2

#not going to make a bar graph for this since it has so few cases

#Double Infections

#make a data frame of just present pathogenes and nymphs of only double co-infections
Pathogen_data_numeric_pain_2_coinfection <- Pathogen_data_numeric_pain %>%
  select(6,7,9,11,12,13)

Pathogen_data_numeric_pain_2_coinfection <- subset(Pathogen_data_numeric_pain_2_coinfection, infection_amount == 2)

#make table which combines all pathogens for double co-infection
pathogen_combinations <- with(Pathogen_data_numeric_pain_2_coinfection,
                              table(`Borrelia burgdorferi sensu lato`, 
                                    `Borrelia miyamotoi `,
                                    `Babesia microti`,
                                    `Anaplasma phagocytophilum`,
                                    `Deer Tick Virus`))

pathogen_combination_df <- as.data.frame(pathogen_combinations)

#bar graph for above data

#re-format data in pathogen_combination_df so it can beasier to convert in a graph
Pathogen_combinations <- c("Bb/Bam","Bb/Ap","Bb/Bmy","Bam/Ap","Bmy/Ap","Ap/DTV")

Pathogen_amounts <- c(40, 15, 4, 2, 1, 1)

pathogen_combinations_and_amounts <- data.frame(Pathogen_combinations, Pathogen_amounts)

coinfection_bar_graph <- ggplot(pathogen_combinations_and_amounts, aes(x = reorder(Pathogen_combinations, -Pathogen_amounts), y = Pathogen_amounts)) +
  geom_bar(stat="identity", fill = "black")

coinfection_bar_graph <- coinfection_bar_graph + theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  panel.grid.major = element_blank(), #remove major gridlines
  panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(fill='transparent') #transparent legend panel
)

coinfection_bar_graph <- coinfection_bar_graph + 
  ggtitle(label = "# of Infected Nymphs for Double Co-infections",
          subtitle = "Bb = Borrelia burgdorferi Bam = Babesia microti Ap = Anaplasmosis phagocytophilum 
          Bmy = Borrelia miyamotoi DTV = Deer Tick Virus") +
  xlab("Pathogen Combination") + 
  ylab("# of I. scapularis Nymphs") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```

```{r co-infection plot, echo=FALSE}
plot(coinfection_bar_graph)
```
Overall, B. burgdorferi and Ba. microti was the common double infection followed by B. burgorferi and A. phagocytophilum. Odd, considering that there was a rougly equal number of infected nymphs for each pathogens. The other pathogen combinations may not be significant given that few nymphs had these specific infections.

Now what we know there is some unique combinations in pathogen infections and the disparity between infection counts between pathogens. Let's look the correlation between these pathogens and enviromental factors

```{r correlation analysis, echo=FALSE, warning=FALSE}
correlation_nymphs <- cor(Pathogen_by_site_spatial_analysisz_no_buffers)

corrplot(correlation_nymphs, method = 'circle', tl.cex = .5)
```
It seems that when we collapse the infection rate of nymphs in sites then almost all pathogens are highly correlated with one another so we need to take care when we try to model how infection rates of a certain pathogen influences pathogen infection of another pathogen. The exception being Powassan virus given that few nymphs were infected, so we won't include it in our analysis. Luckily, the enviromental covariates of TAMES and % forest cover aren't strongly correlated with pathogens, so we can avoid colinearity when modelling. However, there seems to be colinearity in the different types of forest cover so we will only use % forest cover overall ather than dividing into decidious, coniferious, or mixed.

Next let's check out the distributions of pathogen infection by site

```{r histograms, echo=FALSE}

Burgdorferi_infected_nymphs <- Pathogen_by_site_spatial_analysis$total_Borrelia_burgdorferi
miyamotoi_infected_nymphs <- Pathogen_by_site_spatial_analysis$total_Borrelia_miyamotoi
microti_infected_nymphs <- Pathogen_by_site_spatial_analysis$total_Babesia_microti
anaplasma_infected_nymphs <- Pathogen_by_site_spatial_analysis$total_Anaplasma_phagocytophilum
coinfected_nymphs <- Pathogen_by_site_spatial_analysis$total_coinfection_individuals
infected_nymphs <- Pathogen_by_site_spatial_analysis$total_infected_ixodes_scapularis_nymphs

hist(Burgdorferi_infected_nymphs, breaks = "Sturges")
hist(miyamotoi_infected_nymphs)
hist(microti_infected_nymphs)
hist(anaplasma_infected_nymphs)
hist(coinfected_nymphs)
hist(infected_nymphs)

```

Number of infected nymphs at the site level seemed to be skewed at lower values. This could be zero-inflated data but for now we will assume that the distribution follows a gaussian or poisson distribution. At the moment, we will use a binomial model for individual infection and a gaussian distribution for spatial infection when considering sites.

Now that we know infection varies by sites, we need to make maps to visualize that distribution

```{r pressure, echo=FALSE}
Pathogen_by_site_spatial_analysis <- Pathogen_by_site_with_percentages %>%
  select(-c("percent_infected_nymphs", "percent_infection_Bb", "percent_infection_Borrelia_miyamotoi",
            "percent_infection_Borrelia_mayonii", "percent_infection_Babesia_microti","percent_infection_Ehrlichia_muris",
            "percent_infection_Anaplasma_phagocytophilum", "percent_infection_Powassan_DTV", "percent_total_coinfection_individuals",
  ))

Pathogen_by_site_spatial_analysis <- Pathogen_by_site_spatial_analysis %>%
  select(1:25) %>%
  select(-c("Site_Code"))

#only select columns we need
Pathogen_by_site_spatial_analysisz_no_buffers <- Pathogen_by_site_spatial_analysis %>%
  select(1,2,3,4,5,6,7,8,9,10,11,12,13,24)

#replace NA's with zeroes
Pathogen_by_site_spatial_analysisz_no_buffers[is.na(Pathogen_by_site_spatial_analysisz_no_buffers)] <- 0

burgdorferi_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_Borrelia_burgdorferi)) +
  ggtitle("Borrelia Burgdorferi Prevalence Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")

miyamotoi_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_Borrelia_miyamotoi)) +
  ggtitle("Borrelia Miyamotoi Prevalence Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")

microti_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_Babesia_microti)) +
  ggtitle("Babesia Microti Prevalence Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")

anaplasma_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_Anaplasma_phagocytophilum)) +
  ggtitle("Anaplasma pahogyctophilum Prevalence Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")

co_infected_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_coinfection_individuals)) +
  ggtitle("Co-infection Prevalence Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")

infected_map <- ggplot() +
  geom_polygon(data = Rhode_Island, aes(x=long, y = lat, group = group), fill = "grey", alpha =0.8) +
  geom_point(data = Pathogen_by_site_spatial_analysisz_no_buffers,
             aes(x=`Longitude (x) -`, y=`Latitude (y) +`, size=total_infected_ixodes_scapularis_nymphs)) +
  ggtitle("Infected Nymphs Rhode Island") +
  theme(plot.title = element_text(size = 16, face = "bold", color = "black")) +
  labs(size = "# of Infected Nymphs")
```


```{r pressure, echo=FALSE}
plot(burgdorferi_map)
plot(miyamotoi_map)
plot(microti_map)
plot(anaplasma_map)
plot(co_infected_map)
plot(infected_map)
```

The # of infected nymphs don't seem to be connected to any distinct spatial area across all pathogens. This suggests that if there is a spatial correlation it will be for each pathogen rather than across all pathogens. To complement we will be modelling each pathgoen individually to determine the relationships between and among these pathogens.

#Modelling Infection of Tick-borne pathogens in Rhode Island

```{r models, echo=FALSE}
#Below is just simple linear models using total number of infected nymphs by pathogen for each site
#lattitude and longitude are included as covariates to account for spatial variation
#pathogen data is square root transformed to normalize data

#generalized linear model for #of infected nymphs as function of lat long using poisson distribution since data is discrete
mod0 <- glm( sqrt(total_Borrelia_burgdorferi) ~  Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod1 <- glm( sqrt(total_Borrelia_miyamotoi) ~  Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod2 <- glm( sqrt(total_Anaplasma_phagocytophilum) ~  Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod3 <- glm( sqrt(total_Babesia_microti) ~  Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod4 <- glm( sqrt(total_coinfection_individuals) ~ Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod5 <- glm( sqrt(total_coinfection_individuals) ~ Longitude + lattitude + total_Ixodes_Scapularis_nymphs, Pathogen_by_site_final, family = gaussian(link =identity))

#Pathogen prevalence as function of total ixodes scapularis nymphs collected
#here i use a gaussian distribution as the variable being predicted is continious
mod6 <- glm( sqrt(total_Borrelia_burgdorferi) ~ total_Ixodes_Scapularis_nymphs + Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod7 <- glm( sqrt(total_Borrelia_miyamotoi) ~ total_Ixodes_Scapularis_nymphs + Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod8 <- glm( sqrt(total_Anaplasma_phagocytophilum) ~ total_Ixodes_Scapularis_nymphs + Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod9 <- glm( sqrt(total_Babesia_microti) ~ total_Ixodes_Scapularis_nymphs + Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))
mod10 <-glm( sqrt(total_coinfection_individuals) ~ total_Ixodes_Scapularis_nymphs + Longitude + lattitude, Pathogen_by_site_final, family = gaussian(link =identity))

#Negative association between borrelia burgdroferi prevalence such that number of infected nymphs decreases with lattitude
summary(mod0)

#No association with borrelia miyamotoi prevalence with lattitude and longitude
summary(mod1)

#Association with anaplasma phagocytophilum and lattitude such that as lattitude increases, Ap prevalence decreases
summary(mod2)

#Significant relationship with babesia microti such that as longitude increases so does babesia microti prevalence
summary(mod3)

#this model suggests that as ixodes scapularis density increases so to does the number of co-infected nymphs 
summary(mod4)

#co-infections seems to occur more frequently in high density sites regardless of  longitude or lattitude
summary(mod5)

#However, co-nfected indivduals as a proportion of total collected specimens is not significant neither singularly 
#or with lattitude and longitude
summary(mod6)

# no signifigance with percent borrelia burgdorferi infection as function of total IS nymphs with long lat
summary(mod7)

# not signifigance with percent borrelia miyamotoi infection as function of total IS nymphs with long lat
summary(mod8)

#not signifigance with percent babesia microti infection as function of total IS nymphs with long lat
summary(mod9)

#not signifigance with percent anaplasma phagocytohilum infection as function of total IS nymohs
summary(mod10)


```

to normalize the dsitribution for modelling, we square root transformed the # of infected nymphs for each pathogen to make them better suited for linear modelling. Generally, when we include the density of total nymphs found at a location, it seems that only density is significant when determening infection rates such that when density increases it leads to an increase in infection across all pathogens at all sites. The onyl exception is anaplasmosis phagocytophilum which has a significant relationship with lattitude, such that when infection rates decrease when lattitude decreases but with no reciprical relationship with longitude. AIC values when including density are generally lower suggesting these models are a better fit for the data.

Now let's see if B. burgdorferi rates are influenced in anyway by infection rates of other pathogens. For this modelling we will try with all possible combination of pathogens

```{r pathogen regression, echo=FALSE}
zn.lm_1 <- glm(sqrt(total_Borrelia_burgdorferi) ~ sqrt(total_Borrelia_miyamotoi) , Pathogen_by_site_final, family = gaussian(link =identity))


zn.lm_2 <- glm(sqrt(total_Borrelia_burgdorferi) ~ sqrt(total_Anaplasma_phagocytophilum), Pathogen_by_site_final, family = gaussian(link =identity))


zn.lm_3 <- glm( sqrt(total_Borrelia_burgdorferi) ~ sqrt(total_Babesia_microti), Pathogen_by_site_final, family = gaussian(link =identity))

zn.lm_4 <- glm( sqrt(total_Borrelia_burgdorferi) ~ sqrt(total_Babesia_microti)+sqrt(total_Anaplasma_phagocytophilum)+sqrt(total_Borrelia_miyamotoi), Pathogen_by_site_final, family = gaussian(link =identity))

#no signifigance in using borrelia burgdorferi prevalence as a predictor of borrelia miyamotoi prevalence
summary(zn.lm_1)

#Borrelia burgdorferi is signifianct predictor for anaplasma phagocytophilum pathogen prevalence although R-squared is small
summary(zn.lm_2)

#no signifigance in using borrelia burgdorferi prevalence as a predictor of babesia microti prevalence by site
summary(zn.lm_3)

summary(zn.lm_4)
```

It seems that at when pathogens are considered invidiaully as predictors of B. burgdorferi infection rates at the site level they are significant. This isnt suprising considering the earlier we showed that overall infection is strongly tied density. What is interesting is when use all of the pathogens signifigance changes. In this case as demonstrated at the last model we can see that a. phagocytophilum stops being significant when we include ba. microti and b. miyamotoi. This suggests there is a strong relationship bewtween babesia and burgdorferi which we should explore. Looking at the estimate coeffecients we can see that when 1 nymph is infected with burgdorferi it leads an equal increase in the number of microti infected nymphs.

Since looking at this relationship by the site level may generalize across populations rather than individuals we need to look at what effects infection for individual nymphs. Considering the binary response of either infected or not infected. We modelled using binomial distribution using a logit link function in a generalized linear model. We used the covariatesa of TAMES (tick adverse humidity events) and % forest cover with 1 mile of of the site a nymph was collected at. We consider TAMEs alone, TAMEs and forest cover, and the interaction between TAMEs and % forest cover.
```{r models for individuals, echo=FALSE}
#using only the raw data by individuals

#rename columns so we can join
Specimen_Date_Temp_Humidity  <- Specimen_Date_Temp_Humidity %>%
  rename(URI_Vial_Number_on_tube = `VIAL CODE`, Site_Code = `SITE CODE`) 

#modified column so both are characters so join can work  
Specimen_Date_Temp_Humidity$URI_Vial_Number_on_tube <- as.character(Specimen_Date_Temp_Humidity$URI_Vial_Number_on_tube)
  
#left join all the individual data sheets so we can centralize into one data frame
pathogen_data_individual_enviromental_covariates <- Pathogen_data_numeric_pain %>%
  left_join(Specimen_Date_Temp_Humidity, by = join_by(URI_Vial_Number_on_tube)) %>%
  left_join(Sampling_Site_Enviromental_Parameter, by = join_by(Site_Code))

#now only select columns so their are no duplicates
pathogen_data_individual_enviromental_covariates <- pathogen_data_individual_enviromental_covariates %>%
  select(6,7,8,9,10,11,12,13,14,15,25,26,27,38)

#Now we can begin modelling

binomal_mod_1 <- glm(formula = `Borrelia burgdorferi sensu lato`~ TAME ,
            pathogen_data_individual_enviromental_covariates,
    family = binomial(link = "logit"))

binomal_mod_2 <- glm(formula = `Borrelia miyamotoi ` ~ TAME,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_3 <- glm(formula = `Babesia microti` ~ TAME,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_4 <- glm(formula = `Anaplasma phagocytophilum` ~ TAME,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_5 <- glm(formula = Co_infection_present ~ TAME,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_6 <- glm(formula = `Borrelia burgdorferi sensu lato` ~ TAME + `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_7 <- glm(formula = `Borrelia miyamotoi ` ~ TAME + `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_8 <- glm(formula = `Babesia microti` ~ TAME + `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_9 <- glm(formula = `Anaplasma phagocytophilum` ~ TAME + `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_10 <- glm(formula = Co_infection_present ~ TAME + `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_11 <- glm(formula = `Borrelia burgdorferi sensu lato` ~ TAME * `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_12 <- glm(formula = `Borrelia miyamotoi ` ~ TAME * `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_13 <- glm(formula = `Babesia microti` ~ TAME * `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_14 <- glm(formula = `Anaplasma phagocytophilum` ~ TAME * `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

binomal_mod_15 <- glm(formula = Co_infection_present ~ TAME * `% Forest Cover 1 miles`,
            pathogen_data_individual_enviromental_covariates,
            family = binomial(link = "logit"))

summary(binomal_mod_1)
summary(binomal_mod_2)
summary(binomal_mod_3)
summary(binomal_mod_4)
summary(binomal_mod_5)
summary(binomal_mod_6)
summary(binomal_mod_7)
summary(binomal_mod_8)
summary(binomal_mod_9)
summary(binomal_mod_10)
summary(binomal_mod_11)
summary(binomal_mod_12)
summary(binomal_mod_13)
summary(binomal_mod_14)
summary(binomal_mod_15)

```

Based on these models TAMEs are not significant predictors of infection in individual nymphs when considered alone, with an exception for A. phagotyophilum. When considered in conjuction with forest cover, the significance does not change. Additionally, it seems that % forest cover is also not a significant predictor of infection prevalence but this does change when we consider the interaction between TAMES and Forest cover. Conceptually, forest cover and TAMES are proxies for the enviromental and host pressures a nymph may face. For examples, TAMEs represent the amount of time a nymph spends under sub-optimal temperature and humidity events while forest covers is proxy for amount of hosts within the local area. When interactions is conisder TAMES are only significant for B. burgdorferi, forest cover for B. burgdorferi, co-infection, and Ba. microti, and the interaction between them only significant for B. burgdorferi. Partially limiting our conclusions is the low number of infected nymphs we found for b. miyamotoi and co-infected nymphs. AIC scores suggest that the model that includes all enviromental covariates and their interactions is the best fit for this data.

For the next step in this analysis, we will be using parametric statistical testing to determine if pathogens occur at the site or individual levels at greater rates than expected by chance in terms of both abundance or presence.
```{r co-occur code, echo=FALSE, include=FALSE}
#now that know co-infections appear with different pathogen combinations 
#lets see if there is any significance between those pathogen combinations
#make sure to add t() so you can transpose the data frame 
#the specific format for these functions need to site as columns and species as rows


#Subset data set to only pathogen co-occruance within nymphs
Pathogen_for_occur_analysis_by_individual_nymphs <- Pathogen_data_numeric %>%
  select(c(`Borrelia burgdorferi sensu lato`,`Borrelia miyamotoi `,
           `Borrelia mayonii `,`Babesia microti`,`Ehrlichia muris`,`Anaplasma phagocytophilum`,
           `Deer Tick Virus`)) %>%
  rename(`Borrelia burgdorferi` = `Borrelia burgdorferi sensu lato`) %>%
  t()

#subset data to only measure co-occurance within site (not individuals)
#Rename column names for better presentation on graphs
Pathogen_for_occur_analysis_by_sites <- Pathogen_by_site_with_percentages %>%
  select(c(total_Borrelia_burgdorferi ,total_Borrelia_miyamotoi,total_Babesia_microti,
           total_Anaplasma_phagocytophilum,total_Powassan_DTV)) %>%
  rename(`Borrelia burgdorferi` = total_Borrelia_burgdorferi, 
         `Borrelia miyamotoi` = total_Borrelia_miyamotoi,
         `Babesia microti` = total_Babesia_microti,
         ` Anaplasma phagocytophilum`= total_Anaplasma_phagocytophilum,
         `Deer Tick Virus` = total_Powassan_DTV) %>%
  t()

##create objects for for measuring co-occurance of pathogens within seperate dimension


#presence/abscence of pathogens within individuals nymphs
pathogen_occurance_individuals <- cooccur(Pathogen_for_occur_analysis_by_individual_nymphs, spp_names = TRUE)


#prevalence of pathogens by sites

cooccur_pathogen_site_abundance <- cooccur(Pathogen_for_occur_analysis_by_sites, spp_names = TRUE)

##summary information
summary(pathogen_occurance_individuals)
summary(cooccur_pathogen_site_abundance)

#Co-infection between babesia microti and borrelia burgdorferi is significant in co-infection between individuals nymphs

plot(pathogen_occurance_individuals, plotrand = TRUE)

#Presence of pathogens at the site level is signigant between all possible combinations of pathogens except deer tick virus

plot(cooccur_pathogen_site_abundance, plotrand = TRUE)

#presence/abscence of pathogens at sites 
#the code below will change all numbers above zero into 1, this will change it to abscencse/presence rather than abundance
#if you want the code line above this you need to re-run lines 507 - 510 to get the original object back
Pathogen_for_occur_analysis_by_sites[Pathogen_for_occur_analysis_by_sites > 0] <- 1 

pathogen_by_site_presence <- cooccur(Pathogen_for_occur_analysis_by_sites,spp_names = TRUE)

summary(pathogen_by_site_presence)

plot(pathogen_by_site_presence, plotrand = TRUE)

```

```{r co-ccur plots, echo=FALSE}
plot(pathogen_occurance_individuals, plotrand = TRUE)
plot(cooccur_pathogen_site_abundance, plotrand = TRUE)
plot(pathogen_by_site_presence, plotrand = TRUE)
```
The three plots above demonstrate pathogen co-occurance as presence/abscence in individuals , pathogen species abundanceat the site leve, and pathogen species presence at the site level. For individuals the only positive relationship is between B. burgdorferi and Ba. microti, only considering presence. When we consider the site level abundance (# of infected nymphs for each pathogen) there is no non-random association between pathogen abundance. When we consider site presence all combinations (excepting powassan due to limited abundance) are positve non-random associations.

```{r co-ccur plots, echo=FALSE}
summary(pathogen_occurance_individuals)
summary(cooccur_pathogen_site_abundance)
summary(pathogen_by_site_presence)
```

```{r spatial models, echo=FALSE, warning=FALSE}

#these are various models for spatial data
#warning: these are computationally heavy so if you are using a slow computer
#please run them one at a time
Fitticks_miyamotoi_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_miyamotoi, 
                                      x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                      locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                      marginal = zip.gc(link = "log"),
                                      corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")

Fitticks_burgdorferi_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_burgdorferi,
                                        x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                        locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                        marginal = zip.gc(link = "log"),
                                        corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")

Fitticks_microti_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Babesia_microti, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                    locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                    marginal = zip.gc(link = "log"),
                                    corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")

Fitticks_phagocytophilum_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Anaplasma_phagocytophilum, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                            locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                            marginal = zip.gc(link = "log"),
                                            corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")
Fitticks_coinfection_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_coinfection_individuals, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                        locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                        marginal = zip.gc(link = "log"),
                                        corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")

Fitticks_infection_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_infected_ixodes_scapularis_nymphs, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                      locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                      marginal = zip.gc(link = "log"),
                                      corr = matern.gc(kappa = 0.5, nugget =FALSE), method = "GHK")

Fitticks_miyamotoi_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_miyamotoi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                             locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                             marginal = zip.gc(link = "log"),
                                             corr = matern.gc(kappa = 0.5, nugget =TRUE), method = "GHK")

Fitticks_burgdorferi_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_burgdorferi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                               locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                               marginal = zip.gc(link = "log"),
                                               corr = matern.gc(kappa = 0.5, nugget =TRUE), method = "GHK")

Fitticks_microti_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Babesia_microti, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                           locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                           marginal = zip.gc(link = "log"),
                                           corr = matern.gc(kappa = 0.5, nugget =TRUE), method = "GHK")

Fitticks_phagocytophilum_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Anaplasma_phagocytophilum, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                   locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                   marginal = zip.gc(link = "log"),
                                                   corr = matern.gc(kappa = 0.5, nugget =TRUE), method = "GHK")

Fitticks_coinfection_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_coinfection_individuals, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                               locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                               marginal = zip.gc(link = "log"),
                                               corr = matern.gc(kappa = 0.5, nugget =TRUE), method = "GHK")

Fitticks_infection_nugget_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_infected_ixodes_scapularis_nymphs, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                             locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                             marginal = zip.gc(link = "log"),
                                             corr = matern.gc(kappa = 0.5), method = "GHK")

Fitticks_miyamotoi_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_miyamotoi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                 locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                 marginal = zip.gc(link = "log"),
                                                 corr = spherical.gc( nugget =TRUE), method = "GHK")

Fitticks_burgdorferi_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_burgdorferi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                   locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                   marginal = zip.gc(link = "log"),
                                                   corr = spherical.gc( nugget =TRUE), method = "GHK")

Fitticks_microti_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Babesia_microti, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                               locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                               marginal = zip.gc(link = "log"),
                                               corr = spherical.gc(nugget =TRUE), method = "GHK")

Fitticks_phagocytophilum_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Anaplasma_phagocytophilum, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                       locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                       marginal = zip.gc(link = "log"),
                                                       corr = spherical.gc( nugget =TRUE), method = "GHK")

Fitticks_coinfection_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_coinfection_individuals, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                   locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                   marginal = zip.gc(link = "log"),
                                                   corr = spherical.gc( nugget =TRUE), method = "GHK")

Fitticks_infection_nugget_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_infected_ixodes_scapularis_nymphs, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                 locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                 marginal = zip.gc(link = "log"),
                                                 corr = spherical.gc( nugget =TRUE), method = "GHK")

Fitticks_miyamotoi_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_miyamotoi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                          locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                          marginal = zip.gc(link = "log"),
                                          corr = spherical.gc( nugget =FALSE), method = "GHK")

Fitticks_burgdorferi_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Borrelia_burgdorferi, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                            locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                            marginal = zip.gc(link = "log"),
                                            corr = spherical.gc( nugget =FALSE), method = "GHK")

Fitticks_microti_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Babesia_microti, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                        locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                        marginal = zip.gc(link = "log"),
                                        corr = spherical.gc(nugget =FALSE), method = "GHK")

Fitticks_phagocytophilum_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_Anaplasma_phagocytophilum, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                                locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                                marginal = zip.gc(link = "log"),
                                                corr = spherical.gc(nugget =FALSE), method = "GHK")

Fitticks_coinfection_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_coinfection_individuals, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                            locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                            marginal = zip.gc(link = "log"),
                                            corr = spherical.gc(nugget =FALSE), method = "GHK")

Fitticks_infection_sph_nobuffer <- mlegc (y = Pathogen_by_site_spatial_analysisz_no_buffers$total_infected_ixodes_scapularis_nymphs, x = Pathogen_by_site_spatial_analysisz_no_buffers[, 13:14],
                                          locs = Pathogen_by_site_spatial_analysisz_no_buffers[, 1:2], longlat = TRUE,
                                          marginal = zip.gc(link = "log"),
                                          corr = spherical.gc(nugget =FALSE), method = "GHK")
```


```{r pressure, echo=FALSE}

#miyamotoi
summary(Fitticks_miyamotoi_nobuffer)
summary(Fitticks_miyamotoi_nugget_nobuffer)
summary(Fitticks_miyamotoi_sph_nobuffer)
summary(Fitticks_miyamotoi_nugget_sph_nobuffer)

#burgdorferi
summary(Fitticks_burgdorferi_nobuffer)
summary(Fitticks_burgdorferi_nugget_nobuffer)
summary(Fitticks_burgdorferi_sph_nobuffer)
summary(Fitticks_burgdorferi_nugget_sph_nobuffer)

#microti 
summary(Fitticks_microti_nobuffer)
summary(Fitticks_microti_nugget_nobuffer)
summary(Fitticks_microti_sph_nobuffer)
summary(Fitticks_microti_nugget_sph_nobuffer)

#anaplasmosis
summary(Fitticks_phagocytophilum_nobuffer)
summary(Fitticks_phagocytophilum_nugget_nobuffer)
summary(Fitticks_phagocytophilum_sph_nobuffer)
summary(Fitticks_phagocytophilum_nugget_sph_nobuffer)

#coinfection
summary(Fitticks_coinfection_nobuffer)
summary(Fitticks_coinfection_nugget_nobuffer)
summary(Fitticks_coinfection_sph_nobuffer)
summary(Fitticks_coinfection_nugget_sph_nobuffer)

#total infection model

summary(Fitticks_infection_nobuffer)
summary(Fitticks_infection_nugget_nobuffer)
summary(Fitticks_infection_sph_nobuffer)
summary(Fitticks_infection_nugget_sph_nobuffer)

```

While there are pathogens where using a nugget model increased the AIC the change is minor enough that it warrants just considering nugget based models as the default. Even if we where to not consider the nugget by exclusion this would mean it doesn't have a nugget. Adding the nugget parameter just quantifies the effect. In effect the nugget effect quantifies the distance at which local variation of the target variable is so great that no estimate can be made of any spatial association. When the nugget is low it means that a variable is distributed in space beyond randomness while a high nugget means that variability is so great that there no influence from nearby points in space on a spatial location. Only when a nugget is both low relative to the range which is the measure of distance between points and significant can we say there may be a spatial association.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
